% =============================================================================
% ABSTRACT (English mirror of Resumen)
% =============================================================================
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
\markboth{Abstract}{}

\textit{Clutch} performance is one of the most debated concepts in modern
football scouting and one of the least rigorously measured. Public rankings
of the player who shows up when it matters rely on pure event data, with no
tracking, no exogenous counterfactual and---above all---no separation between
the player who genuinely stands out and the player who simply rides the wave
of a team that comes alive. The result is a market that overpays for clutch
reputation without an objective metric. The open question is whether a
player's clutch performance can be estimated causally, net of the collective
pattern of their team, without presupposing that pattern a priori.

This work proposes and evaluates the Expected Clutch Value (\xCV), a causal
framework that estimates the effect of the emotional shock of a goal---both
for and against---on player performance across four channels (offensive,
defensive, off-ball and physical) and three contexts (after scoring, after
conceding and high elimination proximity), residualized against the team's
own delta via a leave-one-out scheme at each shock. The central contribution
of this work is the framework itself; its application to the 2022 FIFA
World Cup with PFF FC data follows from it being the only elite knockout
tournament with open synchronized event and 25\,Hz tracking data, a
condition that enables methodological validation on the most demanding
competitive context without commercial barriers. The output table covers
511~players, 233 with sufficient minutes for reliable analysis
($\geq 270$\,min), and 172 goal-shocks split between 120 from the group
stage and 52 from the knockout rounds.

Causal identification is triangulated through two independent strategies.
The first is a within-player Differences-in-Differences (DiD) estimate over
a $\pm 10$-minute window around each goal, contrasted against modern
estimators for pulse treatments
\parencite{SunAbraham2021,BorusyakJaravelSpiess2024} and with robustness
bounds against parallel-trends violations \parencite{RambachanRoth2023}. The
second is a near-miss quasi-experiment: conditional on comparable pre-shot
$xG$, whether the ball goes in is, in practice, random
\parencite{GauriotPage2019}. The player-level heterogeneous effect is
condensed into a hierarchical multivariate Bayesian Conditional Average
Treatment Effect (CATE) estimated via Hamiltonian Monte Carlo (HMC) with
the No-U-Turn (NUTS) sampler across four chains, with priors informed by
PFF grades and cross-channel correlations modeled through a
Lewandowski--Kurowicka--Joe (LKJ) prior.

The population-average effect is statistically indistinguishable from zero
across the four channels: the shock does not move the average player. The
clutch signal lives in individual heterogeneity and emerges under high
elimination proximity: four players exhibit a significant pressure-clutch
effect at $P(\hat{\beta}>0\mid\text{data}) \geq 0.85$, led by Kylian Mbappé
with a coefficient of $+0.11$ standard deviations and a posterior
probability of $0.97$---consistent with his emergence in the closing stages
of the Argentina-France final. At the channel-specific level, twenty-two
players show a significant effect on the post-scoring offense cell, five
with offensive output rising after scoring and seventeen declining. The
Chasing and Protecting aggregates yield no significant players: the
aggregated effect by shock direction is under-identified at $N = 172$,
whereas the response conditional on the elimination-imminence context is
identifiable. To the best of the present bibliographic review, \xCV stands
as the first framework triangulating within-player DiD with a near-miss
quasi-experiment over a knockout tournament at the player level, across
four channels and residualized against the team block, with honest
quantification of per-player uncertainty.

\vspace{1em}
\noindent\textbf{Keywords:} \xCV; clutch performance; causal identification;
hierarchical Bayesian CATE; PFF FC.

\clearpage
